The Role of Shallow Processing in Agreement Attraction

### 2020-02-23 text/paper_draft.rnw
Updated models for experiment 1 and changed the text location.

source("../scripts/misc.R")

@@

### note: the low accuracy on fillers is not caused by a couple of items
# filler_avgs <- 
# data_exp1 %>% subset(Type == "filler") %>% group_by(Item, condition) %>% 
#               summarize(M = mean(ResponseYes, na.rm = T)) %>%
#               arrange(condition, Item)
# fillers %>% subset(condition == "a") 
# fillers %>% subset(condition == "b")
###

@@

# map to our condition labels
lagoetal_condition_mapping <- data.frame(
    condition = c("a", "b", "c", "d"),
    lagoetal_condition = c("d", "b", "c", "a"), 
    stringsAsFactors = F)

df_lagoetal_attr %<>% left_join( lagoetal_condition_mapping, by = "lagoetal_condition" )

@@

# reformat by-subject averages to a wide format
avg_by_subj_lagoetal_wide <- avg_by_subj_lagoetal %>% 
            mutate(expcond = paste(experiment, condition, sep="_")) %>% 
            ungroup() %>%
            dplyr::select(-experiment, -condition, -avRT, -N,
                          -grammatical, -attractor_num) %>%
            tidyr::spread(expcond, p_yes) %>% 
            mutate(delta_dc = online_d - online_c)
word_freq <- readxl::read_excel("../Data/frequencies_exp1_and_lago.xlsx", sheet = 1)
word_freq$freq_percentage %<>% as.numeric()
word_freq %<>% dplyr::select(-freq_standardized, -freq_percentage, -word)
word_freq %<>% tidyr::spread(place, freq_count)

word_freq %<>% dplyr::group_by(exp) %>% 
               dplyr::mutate( cat_n1 = ifelse(n1 > median(n1), "high", "low"),
                              cat_n2 = ifelse(n2 > median(n2), "high", "low"),
                              freqlog_n1 = scale(log(n1)),
                              freqlog_n2 = scale(log(n2)),
                              freqlog_n1n2 = log(n1) - log(n2)
                            ) %>% 
                ungroup()
word_freq$cat_n1 %<>% as.factor()
word_freq$cat_n2 %<>% as.factor()

word_freq_exp1 <- word_freq %>% 
                    subset(exp == "up") %>% 
                    dplyr::select(-exp)
word_freq_lago <- word_freq %>% 
                    subset(exp == "lago") %>% 
                    dplyr::select(-exp)

@@

# identify bad participants 
bad_subjects_lagoetal <- subset(avg_by_subj_lagoetal_wide, delta_dc <= 0.25 ) %>% .$subject
df_lagoetal_attr_clean <- df_lagoetal_attr %>% subset(!subject %in% bad_subjects_lagoetal)
df_merge_exp1 <- data_exp1_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, 
                                    grammatical, attractor_num,
                                    # condition,
                                    subject, item=Item,
                                    ResponseYes, RT)
df_merge_exp1$experiment <- "Experiment 1"
df_merge_exp1$grammatical <- with(df_merge_exp1, 
                                  ifelse(grammatical == "gram", 
                                          "grammatical",
                                          "ungrammatical"))
df_merge_exp1$attractor_num <- with(df_merge_exp1, 
                                    ifelse(attractor_num == "pl",
                                           "plural", 
                                           "singular"))
df_merge_exp1$item %<>% as.integer()
df_merge_exp1$subject %<>% as.character()
#df_merge_exp1 %<>% left_join(word_freq_exp1, by = "item")

df_merge_lago <- df_lagoetal_attr_clean %>%
                      ungroup() %>% 
                      dplyr::select(grammatical, attractor_num,
                                    subject, item, ResponseYes, RT)
#df_merge_lago %<>% left_join(word_freq_lago, by = "item")
df_merge_lago$experiment <- "Lago et al. (2018)" 
df_merge_lago$source <- NA
df_merge_lago$item %<>% add(1000)

@@

Prior to analysis we removed the data for all participants who failed to show sufficient sentitivity to the effect of grammaticality in singular attractor conditions, i.e., when no agreement attraction was expected. Specifically, we removed all participants for whom the difference in the percentage of `yes' responses between the grammatical condition \ref{item:exp1expitem-sgsg} and the ungrammatical condition \ref{item:exp1expitem-sgpl} was below the threshold of 25 percentage points. As a result, we excluded \Sexpr{length(bad_subjects)} participants from experiment 1, and \Sexpr{stopifnot(length(bad_subjects_lagoetal) == 1)} one participant from the Lago et al. experiment.

@@

fname_exp1_responses_exp1only <- "../workspace_exp1/fit_responses_exp1only"
m_responses_exp1only <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) + 
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers %>% subset(experiment == "Experiment 1"),
                   family = bernoulli("probit"), 
                   chains = n_chains, cores = n_cores, iter = n_iter,
                   file = fname_exp1_responses_exp1only)

fname_exp1_responses_Lagoonly <- "../workspace_exp1/fit_responses_Lagoonly"
m_responses_Lagoonly <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) + 
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers %>% subset(experiment != "Experiment 1"),
                   family = bernoulli("probit"), 
                   chains = n_chains, cores = n_cores, iter = n_iter,
                   file = fname_exp1_responses_Lagoonly)

fname_exp1_rts <- "../workspace_exp1/fit_responses_rt"
m_rts <- brm(RT ~ cEndsInConsonant * cUngrammatical * cAttractorPlural +
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers,
                   family = lognormal(),
                   chains = n_chains, cores = n_cores, iter = n_iter,
                   file = fname_exp1_rts)

@@

suppressWarnings({
print(p_m_response + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment"))
})

@@

<<exp1ResponseModelSepExp1, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Experiment 1.', fig.height=1.5, fig.align='center'>>=
contrast_names_exp1 <- c("cUngrammatical" = "Experiment 1: Ungrammaticality",
                         "cAttractorPlural" = "Experiment 1: Plural Attactor",
                         "cUngrammatical:cAttractorPlural" = "Experiment 1: Ungrammaticality * Plural Attractor"
                        )

@@

suppressWarnings({
print(p_m_response_exp1 + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment"))
})

contrast_names_lago <- c("cUngrammatical" = "Lago et al.: Ungrammaticality",
                         "cAttractorPlural" = "Lago et al.: Plural Attactor",
                         "cUngrammatical:cAttractorPlural" = "Lago et al.: Ungrammaticality * Plural Attractor"
                        )

<<exp1ResponseModelSepExp2, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Lago et al. (2018).', fig.height=1.5, fig.align='center'>>=
p_m_response_lago <- 
  create_model_coefs_plot( m_responses_Lagoonly, 
        plot_stats = T, map_names = contrast_names_lago,
        expand_right = 2.5, expand_top = 2, x_stat_adjust = 1.1,
        x_breaks = c(-3,-2,-1, 0,1, 2) ) + 
        xlab("Estimate (probit)")

suppressWarnings({
print(p_m_response_lago + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment"))
})

>>>>

# to-do: check that the CI-computation code is up-to-date
source("../scripts/misc.R")

@@

### note: the low accuracy on fillers is not caused by a couple of items
###       -> see below
# filler_avgs <- 
# data_exp1 %>% subset(Type == "filler") %>% group_by(Item, condition) %>% 
#               summarize(M = mean(ResponseYes, na.rm = T)) %>%
#               arrange(condition, Item)
# fillers %>% subset(condition == "a") 
# fillers %>% subset(condition == "b")
###

@@

# map to our condition labels
lagoetal_condition_mapping <- data.frame( condition = c("a", "b", "c", "d"),
                                          lagoetal_condition = c("d", "b", "c", "a"), 
                                          stringsAsFactors = F)
df_lagoetal_attr %<>% left_join( lagoetal_condition_mapping, by = "lagoetal_condition" )

@@

# reformat by-subject averages to a wide format
avg_by_subj_lagoetal_wide <- avg_by_subj_lagoetal %>% 
                                    mutate(expcond = paste(experiment, condition, sep="_")) %>% 
                                    ungroup() %>%
                                    dplyr::select(-experiment, -condition, -avRT, -N,
                                                  -grammatical, -attractor_num) %>%
                                    tidyr::spread(expcond, p_yes) %>% 
                                    mutate(delta_dc = online_d - online_c)

@@

# identify bad participants 
bad_subjects_lagoetal <- subset(avg_by_subj_lagoetal_wide, delta_dc <= 0.25 ) %>% .$subject
df_lagoetal_attr_clean <- df_lagoetal_attr %>% subset(!subject %in% bad_subjects_lagoetal)

# merge both datasets
df_merge_exp1 <- data_exp1_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, 
                                    grammatical, 
                                    attractor_num,
                                    # condition,
                                    subject, 
                                    item=Item,
                                    ResponseYes, 
                                    RT)
df_merge_exp1$experiment <- "Experiment 1"
df_merge_exp1$grammatical %<>% dplyr::recode(gram="grammatical", ungram="ungrammatical")
df_merge_exp1$attractor_num %<>% dplyr::recode(pl="plural", sg="singular")

df_merge_exp1$item %<>% as.integer()
df_merge_exp1$subject %<>% as.character()

df_merge_lago <- df_lagoetal_attr_clean %>%
                      ungroup() %>% 
                      dplyr::select(grammatical, attractor_num,
                                    subject, item, ResponseYes, RT)
df_merge_lago$experiment <- "Lago et al. (2018)" 
df_merge_lago$source <- NA

word_freq <- readxl::read_excel("../Data/frequencies_exp1_and_lago.xlsx", sheet = 1)
word_freq$freq_percentage %<>% as.numeric()
word_freq %<>% dplyr::select(-freq_standardized, -freq_percentage, -word)
word_freq %<>% tidyr::spread(place, freq_count)

word_freq %<>% dplyr::group_by(exp) %>% 
               dplyr::mutate( freq_cat_n1 = ifelse(n1 > median(n1), "high", "low"),
                              freq_cat_n2 = ifelse(n2 > median(n2), "high", "low"),
                              freqlog_n1 = scale(log(n1)),
                              freqlog_n2 = scale(log(n2)),
                              freqlog_n1n2 = log(n1) - log(n2)
                            ) %>% 
                ungroup()
# word_freq$freq_cat_n1 %<>% as.factor()
# word_freq$freq_cat_n2 %<>% as.factor()

word_freq_exp1 <- word_freq %>% subset(exp == "exp1") %>% dplyr::select(-exp)
df_merge_exp1 %<>% left_join(word_freq_exp1, by = "item")

word_freq_lago <- word_freq %>% subset(exp == "lagoetal") %>% dplyr::select(-exp)
df_merge_lago %<>% left_join(word_freq_lago, by = "item")
df_merge_lago$item %<>% add(1000)

@@

Prior to analysis we removed the data for all participants who failed to show sufficient sentitivity to the effect of grammaticality in singular attractor conditions, i.e., when no agreement attraction was expected. Specifically, we removed all participants for whom the difference in the percentage of `yes' responses between the grammatical condition \ref{item:exp1expitem-sgsg} and the ungrammatical condition \ref{item:exp1expitem-sgpl} was below the threshold of 25 percentage points. As a result, we excluded \Sexpr{length(bad_subjects)} participants from experiment 1, and \Sexpr{stopifnot(length(bad_subjects_lagoetal) == 1)} one participant from the Lago et al. data

@@

fname_exp1_responses_exp1only <- "../workspace_exp1/fit_responses_exp1only"
m_responses_exp1only <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural * freqlog_n1n2 + 
                                 (cUngrammatical * cAttractorPlural * freqlog_n1n2 + 1| subject) + 
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers %>% subset(experiment == "Experiment 1"),
                   family = bernoulli("probit"), 
                   chains = n_chains, cores = n_cores, iter = n_iter, init_r = .1,
                   file = fname_exp1_responses_exp1only
                   )

fname_exp1_responses_Lagoonly <- "../workspace_exp1/fit_responses_lagoonly"
m_responses_Lagoonly <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural * freqlog_n1n2 + 
                                 (cUngrammatical * cAttractorPlural * freqlog_n1n2 + 1| subject) + 
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers %>% subset(experiment != "Experiment 1"),
                   family = bernoulli("probit"), 
                   chains = n_chains, cores = n_cores, iter = n_iter, init_r = .1,
                   file = fname_exp1_responses_Lagoonly)

fname_exp1_rts <- "../workspace_exp1/fit_responses_rt"
m_rts <- brm(RT ~ cEndsInConsonant * cUngrammatical * cAttractorPlural * freqlog_n1n2 +
                                 (cUngrammatical * cAttractorPlural * freqlog_n1n2 + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers,
                   family = lognormal(),
                   chains = n_chains, cores = n_cores, iter = n_iter,
                   file = fname_exp1_rts)

@@

suppressWarnings({
print(p_m_response + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment") + facet_wrap(~"Experiment 1 and Lago et al. (2018)"))
})

@@

<<exp1ResponseModelSepExp1AndLago, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Experiment1 and Lago et al. (2018).', fig.height=5, fig.align='center'>>=

contrast_names_exp1 <- c("cUngrammatical" = "Ungrammaticality",
                         "cAttractorPlural" = "Plural Attactor",
                         "freqlog_n1n2"="log Frequency: N1-N2",
                         "cUngrammatical:freqlog_n1n2" = "Ungrammaticality * log Frequency: N1-N2",
                         "cAttractorPlural:freqlog_n1n2" = "Plural Attactor * log Frequency: N1-N2",
                         "cUngrammatical:cAttractorPlural" = "Ungrammaticality * Plural Attractor",
                         "cUngrammatical:cAttractorPlural:freqlog_n1n2" = "Ungrammaticality * Plural Attractor * log Frequency: N1-N2"
                        )

@@

p_m_response_exp1 <- p_m_response_exp1 + facet_wrap(~"Experiment 1") + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment")

p_m_response_lago <- 
  create_model_coefs_plot( m_responses_Lagoonly, 
        plot_stats = T, map_names = contrast_names_exp1,
        expand_right = 2.5, expand_top = 2, x_stat_adjust = 1.1,
        x_breaks = c(-3,-2,-1, 0,1, 2) ) + 
        xlab("Estimate (probit)")

p_m_response_lago <- p_m_response_lago + facet_wrap(~"Lago et al. (2018)") + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment")

p_m_response_exp1_and_lago <- ggarrange(p_m_response_exp1, p_m_response_lago, ncol = 1)

suppressWarnings({
print(p_m_response_exp1_and_lago)
})
